Implementing XBF: An Efficient Self-Describing Binary Format

### 2023-08-18 docs/presentation/presentation.tex
Update paper and presentation

% \setbeameroption{show notes}

@@

\begin{frame}{Shortcomings of Current Formats}
	\begin{itemize}
		\item Repeated sending of metadata.
		\item Inefficiency of plain text.
		      \note[item]{Expensive conversion between plain text and binary}
		      \note[item]{Requires more bytes to store the same data}
		\item Use of big-endian.
		      \note[item]{Most modern processors (x86 and ARM) are little-endian}
	\end{itemize}
\end{frame}

\begin{frame}{Improvements For a New Format}
	\begin{itemize}
		\item When possible, send metadata only a single time (mainly for homogenous lists).
		\item Allow sending metadata to be optional, such as if a client already has it.
		      \note[item]{Keep metadata and data separate, not coupled together}
		\item Utilize little-endian to reduce conversions.
		\item Prioritize a simple type system.
		      \note[item]{Hopefully makes the implementation simpler}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Primitives}
	\begin{itemize}
		\item Boolean
		      \note[item]{True of False, one byte}
		      \note[item]{All integers are little-endian}
		      \note[item]{Integer types are equivalent to the types in stdint.h C header}
		\item U8, U16, U32, U64, U128, U256
		      \note[item]{Unsigned Integer Types}
		      \begin{itemize}
			      \item Equivalent to uint*\_t types in C
		      \end{itemize}
		\item I8, I16, I32, I64, I128, I256
		      \note[item]{Signed Integer Types}
		      \begin{itemize}
			      \item Equivalent to int*\_t types in C
		      \end{itemize}
		\item F32, F64
		      \note[item]{Floating Point Types}
		      \begin{itemize}
			      \item IEEE 754 floating point numbers
			      \item Equivalent to float and double in C
		      \end{itemize}
		\item Bytes  (Non-UTF-8 Bytes)
		      \note[item]{String but doesn't have to be UTF-8}
		\item String (UTF-8 Bytes)
		      \note[item]{Sends length as u64 followed by UTF-8 bytes}
	\end{itemize}
\end{frame}

\begin{frame}{Primitive Metadata}
	\center
	\begin{tabular}{|c|c|}
		\hline
		\textbf{Type} & \textbf{Discriminant} \\
		\hline
		Boolean       & 0                     \\
		U8            & 1                     \\
		U16           & 2                     \\
		U32           & 3                     \\
		...           & ...                   \\
		I256          & 12                    \\
		F32           & 13                    \\
		F64           & 14                    \\
		Bytes         & 15                    \\
		String        & 16                    \\
		\hline
	\end{tabular}
	\note[item]{Metadata is sent as a single byte discriminant value}
	\note[item]{The discriminant begin at 0 with Boolean, and increase by 1 for each type}
	\note[item]{Progress in the same order as the previous slide, with String being the final value}
\end{frame}

@@

				DY = Discriminant of field "last"}
			\center
			\begin{bytefield}{32}
				\bitbox{2}{\rot{DS}} & \bitbox{16}{struct name length} & \bitboxes{2}{Point} & \bitbox{4}{fields} \\
				\bitboxes{2}{{18} {5} {0} {0} {0} {0} {0} {0} {0} {80} {\rot{111}} {\rot{105}} {\rot{110}} {\rot{116}} {2} {0}} \\
			\end{bytefield}
			\begin{bytefield}{32}
				\bitbox{16}{first name length} & \bitboxes{2}{first} & \bitbox{2}{\rot{DX}} \\
				\bitboxes{2}{{5} {0} {0} {0} {0} {0} {0} {0} {\rot{102}} {\rot{105}} {\rot{114}} {\rot{115}} {\rot{116}} {7}} \\
			\end{bytefield}
			\begin{bytefield}{32}
				\bitbox{16}{last name length} & \bitboxes{2}{last} & \bitbox{2}{\rot{DY}} \\
				\bitboxes{2}{{4} {0} {0} {0} {0} {0} {0} {0} {\rot{108}} {\rot{97}} {\rot{115}} {\rot{116}} {7}}
			\end{bytefield}
		\end{column}
		\begin{column}{0.35\textwidth}
			\begin{lstlisting}
struct Point {
	first: i8,
	last: i8
}
\end{lstlisting}
		\end{column}
	\end{columns}
\end{frame}

>>>>

\setbeameroption{show notes}

@@

\begin{frame}{Shortcomings of Current Formats}
	\begin{itemize}
		\item Repeated sending of metadata.
		\item Inefficiency of plain text.
		      \note[item]{Expensive conversion between plain text and binary}
		      \note[item]{Requires more bytes to store the same data}
		\item Use of big-endian.
		      \note[item]{Most modern processors (x86 and ARM) are little-endian}
		      \note[item]{Both CBOR and MessagePack are big-endian}
	\end{itemize}
	% XML from paper here
\end{frame}

% possibly add three slides explaining each of these
% 1: XML being huge
% 2: plain text example from paper
% 3: 1027 as u32 0 4 3 0 -> 0 3 4 0

\begin{frame}{Improvements For a New Format}
	\begin{itemize}
		\item When possible, send metadata only a single time (mainly for homogenous lists).
		\item Allow sending metadata to be optional, such as if a client already has it.
		      \note[item]{Keep metadata and data separate, not coupled together}
		\item Utilize little-endian to eliminate endian conversions on common architectures.
		\item Prioritize a simple type system.
		      \note[item]{Hopefully makes the implementation simpler}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Primitives}
	\begin{itemize}
		\item Boolean
		      \note[item]{True of False, one byte}
		      \note[item]{All integers are little-endian}
		      \note[item]{Integer types are equivalent to the types in stdint.h C headers}
		\item U8, U16, U32, U64, U128, U256
		      \note[item]{Unsigned Integer Types}
		      \begin{itemize}
			      \item Equivalent to uint*\_t types in C
		      \end{itemize}
		\item I8, I16, I32, I64, I128, I256
		      \note[item]{Signed Integer Types}
		      \begin{itemize}
			      \item Equivalent to int*\_t types in C
		      \end{itemize}
		\item F32, F64
		      \note[item]{Floating Point Types}
		      \begin{itemize}
			      \item IEEE 754 floating point numbers
			      \item Equivalent to float and double in C
		      \end{itemize}
		\item Bytes  (Non-UTF-8 Bytes)
		      \note[item]{String but doesn't have to be UTF-8}
		\item String (UTF-8 Bytes)
		      \note[item]{Sends length as u64 followed by UTF-8 bytes}
	\end{itemize}
\end{frame}

\begin{frame}{Primitive Metadata}
	\center
	\begin{tabular}{|c|c|}
		\hline
		\textbf{Type} & \textbf{Discriminant} \\
		\hline
		Boolean       & 0                     \\
		U8            & 1                     \\
		U16           & 2                     \\
		U32           & 3                     \\
		...           & ...                   \\
		I256          & 12                    \\
		F32           & 13                    \\
		F64           & 14                    \\
		Bytes         & 15                    \\
		String        & 16                    \\
		\hline
	\end{tabular}
	% two columsn for each half of the list
	\note[item]{Metadata is sent as a single byte discriminant value}
	\note[item]{The discriminant begin at 0 with Boolean, and increase by 1 for each type}
	\note[item]{Progress in the same order as the previous slide, with String being the final value}
\end{frame}

@@

				DY = Discriminant of field "last"}
			\center
			\begin{bytefield}{32}
				\bitbox{2}{\rot{DS}} & \bitbox{16}{struct name length} & \bitboxes{2}{Point} & \bitbox{4}{fields} \\
				\bitboxes{2}{{18} {5} {0} {0} {0} {0} {0} {0} {0} {80} {\rot{111}} {\rot{105}} {\rot{110}} {\rot{116}} {2} {0}} \\
			\end{bytefield}
			\begin{bytefield}{32}
				\bitbox{16}{first name length} & \bitboxes{2}{first} & \bitbox{2}{\rot{DX}} \\
				\bitboxes{2}{{5} {0} {0} {0} {0} {0} {0} {0} {\rot{102}} {\rot{105}} {\rot{114}} {\rot{115}} {\rot{116}} {7}} \\
			\end{bytefield}
			\begin{bytefield}{32}
				\bitbox{16}{last name length} & \bitboxes{2}{last} & \bitbox{2}{\rot{DY}} \\
				\bitboxes{2}{{4} {0} {0} {0} {0} {0} {0} {0} {\rot{108}} {\rot{97}} {\rot{115}} {\rot{116}} {7}}
			\end{bytefield}
		\end{column}
		\begin{column}{0.35\textwidth}
			\begin{lstlisting}
struct Point {
	first: i8,
	last: i8
}
\end{lstlisting}
		\end{column}
	\end{columns}
	% add explanation for DS, DX, DY
\end{frame}
